---
title: "For Meeting on Nov 20"
author: Jay Kim
output: github_document
pdf_document: default
editor_options:
  chunk_output_type: inline
---
```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(ggalt)
data <- read_csv("/Users/jaykim/Documents/EDA/eda20-team5-project/Jay's work/data.csv")
```

```{r}
head(data)
```

# Analysis 1: College enrollment in nearby institutions  
Percentage 18-24 year-olds enrolled in college within 25-mile radius.

LOS ANGELES
```{r}
#4 diff datasets each for 4 diff regions
college_enrolled_NYC <- data%>%
  filter(msaname15 == "New York-Newark-Jersey City, NY-NJ-PA Metro Area")

college_enrolled_LA <- data%>%
  filter(msaname15 == "Los Angeles-Long Beach-Anaheim, CA Metro Area")

college_enrolled_Hou <- data%>%
  filter(msaname15 == "Houston-The Woodlands-Sugar Land, TX Metro Area")

college_enrolled_Philly <- data%>%
  filter(msaname15 == "Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area")

```

```{r}
ggplot() +
  geom_density(aes(ED_COLLEGE), alpha = .2, col="blue", data = college_enrolled_NYC) +
  geom_density(aes(ED_COLLEGE),  alpha = .2, col="darkgreen",  data = college_enrolled_LA)+
  geom_density(aes(ED_COLLEGE),  alpha = .2, col="purple",  data = college_enrolled_Hou)+
  geom_density(aes(ED_COLLEGE),  alpha = .2,  col="red", data = college_enrolled_Philly)+
    theme_classic()+ 
  theme(panel.grid.major.x=element_line())+ #add grid lines
  scale_x_continuous(
            breaks=seq(20, 70, 5),
            labels = function(x){paste0(x*1, "%")}
            )+ #change x axis labels+
  labs(title = "Southern California High School's College enrollment rates in nearby institutions",
       x="Percent"
       )
```




# Analysis 2: see how  HS grad rates changed from 2010 to 2015
Los Angeles
- Los Angeles-Long Beach-Anaheim, CA Metro Area
- Riverside-San Bernardino-Ontario, CA Metro Area	
- San Diego-Carlsbad, CA Metro Area	
- Oxnard-Thousand Oaks-Ventura, CA Metro Area	

New York
```{r}
head(data)
NYC <- data %>%
  filter(stateusps == "NY")%>%
  group_by(msaname15)%>%
  summarise(count=n())%>%
  arrange(desc(count))

NYC
```

Houston
```{r}
head(data)
TX <- data %>%
  filter(stateusps == "TX")%>%
  group_by(msaname15)%>%
  summarise(count=n())%>%
  arrange(desc(count))

TX
```

Philadelphia 
```{r}
head(data)
PA <- data %>%
  filter(stateusps == "PA")%>%
  group_by(msaname15)%>%
  summarise(count=n())%>%
  arrange(desc(count))

PA
```


data manip:
```{r}
data5 <-data %>% 
  drop_na(ED_HSGRAD) #getting rid of NA values

#average grad rates in 2015
grad_rates_15<- data5%>%
  filter(msaname15 %in% c("Riverside-San Bernardino-Ontario, CA Metro Area", "San Diego-Carlsbad, CA Metro Area","Oxnard-Thousand Oaks-Ventura, CA Metro Area","Los Angeles-Long Beach-Anaheim, CA Metro Area" ,"Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area", "Pittsburgh, PA Metro Area","Scranton--Wilkes-Barre--Hazleton, PA Metro Area","Houston-The Woodlands-Sugar Land, TX Metro Area	","New York-Newark-Jersey City, NY-NJ-PA Metro Area	"))%>%
  filter(year == '2015')%>%
  group_by(msaname15, year)%>%
  summarise(mean_grad_2015 = mean(ED_HSGRAD))%>%
  mutate(region = sapply(strsplit(msaname15, ","),"[", 1))
#grad_rates_15


#average grad rates in 2010
grad_rates_10<- data5%>%
  filter(msaname15 %in% c("Riverside-San Bernardino-Ontario, CA Metro Area", "San Diego-Carlsbad, CA Metro Area","Oxnard-Thousand Oaks-Ventura, CA Metro Area","Los Angeles-Long Beach-Anaheim, CA Metro Area" ,"Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area", "Pittsburgh, PA Metro Area","Scranton--Wilkes-Barre--Hazleton, PA Metro Area","Houston-The Woodlands-Sugar Land, TX Metro Area	","New York-Newark-Jersey City, NY-NJ-PA Metro Area	"))%>%
  filter(year == '2010')%>%
  group_by(msaname15, year)%>%
  summarise(mean_grad_2010 = mean(ED_HSGRAD))%>%
  mutate(region = sapply(strsplit(msaname15, ","),"[", 1))
#grad_rates_10

#Join on msaname15
grad_rate<- inner_join(grad_rates_10,grad_rates_15, by= "region")%>%
  select(region, year.x, year.y, mean_grad_2010, mean_grad_2015) #selecting relevant columns
#grad_rate 
```

plot: 
```{r}
ggplot(grad_rate, aes(y=region, x = mean_grad_2015, xend = mean_grad_2010)) + 
        geom_dumbbell(color= "olivedrab", 
                      size=0.75, 
                      point.colour.l="black") +
  labs(title = "Change in high school graduation rate", 
       x = "Percent Graduated",
       y = "Region")+
  theme_classic()+
  theme(panel.grid.major.x=element_line())+
   scale_x_continuous(
            breaks=seq(76, 100, 1),
            labels = function(x){paste0(x*1, '%')}
            )

#or
#   scale_x_continuous(
#            breaks=seq(76, 82, 0.5),
#            labels = function(x){paste0(x*1, '%')}
#            )
  
#how to indicate which change is positive vs negative
#ifelse(mean_grad_2015 > mean_grad_2010, "red", "blue" : not working
#idk how to drop "Metro Area"
#how to add percent symbols 
```






# Analysis 3: AP Classes x Single-Headed Households
Data manipulation:

LA
```{r}
data1 <-data %>% 
  drop_na(ED_APENR,SE_SINGLE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

LA <- data1%>%
  filter(msaname15 =="Los Angeles-Long Beach-Anaheim, CA Metro Area")%>% #looking at only LA
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), avg_single = mean(SE_SINGLE))%>% 
  #column called "avg_AP" that is avg % of students in AP
  #column called "avg_single" that is average % of single headed households
  arrange(desc(avg_AP))
#LA 

#compare 'CA' with % of AP enrolled after u take out cities that have % of single-headed household >30
LA_less_single <- LA%>%
  filter(avg_single<15)%>%
  arrange(desc(avg_single))
#LA_less_single
```

Houston 
```{r}
data1 <-data %>% 
  drop_na(ED_APENR, SE_SINGLE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

Hou <- data1%>%
  filter(msaname15 =="Houston-The Woodlands-Sugar Land, TX Metro Area")%>% #looking at only Houston
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), avg_single = mean(SE_SINGLE))%>% 
  #column called "avg_AP" that is avg % of students in AP
  #column called "avg_single" that is average % of single headed households
  arrange(desc(avg_AP))
#Hou 

#compare 'Hou' with % of AP enrolled after u take out cities that have % of single-headed household >30
Hou_less_single <- Hou%>%
  filter(avg_single<15)%>%
  arrange(desc(avg_single))
#Hou_less_single
```

NYC 
```{r}
data1 <-data %>% 
  drop_na(ED_APENR,SE_SINGLE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

NYC <- data1%>%
  filter(msaname15 =="New York-Newark-Jersey City, NY-NJ-PA Metro Area")%>% #looking at only NYC
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), avg_single = mean(SE_SINGLE))%>% 
  #column called "avg_AP" that is avg % of students in AP
  #column called "avg_single" that is average % of single headed households
  arrange(desc(avg_AP))
#NYC 

#compare 'NY' with % of AP enrolled after u take out cities that have % of single-headed household >30
NYC_less_single <- NYC%>%
  filter(avg_single<15)%>%
  arrange(desc(avg_single))
#NYC_less_single
```

Philly
```{r}
data1 <-data %>% 
  drop_na(ED_APENR, SE_SINGLE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

Philly <- data1%>%
  filter(msaname15 =="Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area")%>% #looking at only Houston
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), avg_single = mean(SE_SINGLE))%>% 
  #column called "avg_AP" that is avg % of students in AP
  #column called "avg_single" that is average % of single headed households
  arrange(desc(avg_AP))
#Philly 

#compare 'Hou' with % of AP enrolled after u take out cities that have % of single-headed household >30
Philly_less_single <- Philly%>%
  filter(avg_single<15)%>%
  arrange(desc(avg_single))
#Philly_less_single
```


Plots: 
```{r}
LA_AP<- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = LA)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = LA_less_single)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Los Angeles",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
LA_AP

NYC_AP <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = NYC)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = NYC_less_single)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="New York, Newark, Jersey City",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
NYC_AP


Hou_AP <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = Hou)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = Hou_less_single)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Houston",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
Hou_AP

Philly_AP <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = Philly)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = Philly_less_single)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Philadelphia-Camden-Wilmington",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
Philly_AP
```


In the US, the two most common ways of measuring academic abilities include GPA and Standardized Test Scores. While we did not have those variables to analyze, we did have the ratio of 11th/12th grades enrolled in at least 1 AP class to the total number of students at that high school. To make the best out of the data we were presented, we decided to analyze how home life affects a high school student's ambition in the classroom. Across all four major cities, we see a general trend that more students at schools where less than 15% of the student's families are single-headed enrolled in at least 1 AP class. 
Another home-life factor there was to analyze was median household income. 


# Analysis 4: AP Classes x Median household income
Data manipulation:

LA
```{r}
data1 <-data %>% 
  drop_na(ED_APENR,SE_MHE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

LA <- data1%>%
  filter(msaname15 =="Los Angeles-Long Beach-Anaheim, CA Metro Area")%>% #looking at only LA
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), mean_inc = mean(SE_MHE))%>% 
  #column called "avg_AP" that is avg % of students in AP
  #column called "avg_median" that is median income
  arrange(desc(avg_AP))
#LA 

#compare 'CA' with % of AP enrolled after u take out cities that have % of single-headed household >30
LA_wealthy <- LA%>%
  filter(mean_inc>90000)%>%
  arrange(desc(mean_inc))
#LA_wealthy
```

Houston 
```{r}
data1 <-data %>% 
  drop_na(ED_APENR, SE_MHE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

Hou <- data1%>%
  filter(msaname15 =="Houston-The Woodlands-Sugar Land, TX Metro Area")%>% #looking at only Houston
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), mean_inc = mean(SE_MHE))%>% 
  #column called "avg_AP" that is avg % of students in AP
   #column called "avg_median" that is median income
  arrange(desc(avg_AP))
#Hou 

#compare 'Hou' with % of AP enrolled after u take out cities that have % of single-headed household >30
Hou_wealthy <- Hou%>%
  filter(mean_inc>90000)%>%
  arrange(desc(mean_inc))
#Hou_wealthy
```

NYC 
```{r}
data1 <-data %>% 
  drop_na(ED_APENR,SE_MHE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

NYC <- data1%>%
  filter(msaname15 =="New York-Newark-Jersey City, NY-NJ-PA Metro Area")%>% #looking at only NYC
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), mean_inc = mean(SE_MHE))%>% 
  #column called "avg_AP" that is avg % of students in AP
   #column called "avg_median" that is median income
  arrange(desc(avg_AP))
#NYC 

#compare 'NY' with % of AP enrolled after u take out cities that have % of single-headed household >30
NYC_wealthy <- NYC%>%
  filter(mean_inc>90000)%>%
  arrange(desc(mean_inc))
#NYC_wealthy
```

Philly
```{r}
data1 <-data %>% 
  drop_na(ED_APENR, SE_MHE) #getting rid of NA values
#ED_APENR : Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.

Philly <- data1%>%
  filter(msaname15 =="Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area")%>% #looking at only Houston
  group_by(geoid)%>%  #group by cities/counties in CA
  summarise(avg_AP= mean(ED_APENR), mean_inc = mean(SE_MHE))%>% 
  #column called "avg_AP" that is avg % of students in AP
   #column called "avg_median" that is median income
  arrange(desc(avg_AP))
#Philly 

#compare 'Hou' with % of AP enrolled after u take out cities that have % of single-headed household >30
Philly_wealthy <- Philly%>%
  filter(mean_inc>90000)%>%
  arrange(desc(mean_inc))
#Philly_wealthy
```


Plots: 
```{r}
LA_AP<- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = LA)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = LA_wealthy)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Los Angeles",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
LA_AP

NYC_AP <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = NYC)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = NYC_wealthy)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="New York, Newark, Jersey City",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
NYC_AP


Hou_AP <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = Hou)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = Hou_wealthy)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Houston",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
Hou_AP

Philly_ap <- ggplot()+
  geom_density(aes(avg_AP), alpha = .2, fill="pink", data = Philly)+
  geom_density(aes(avg_AP), alpha = .2, fill="lightblue",  data = Philly_wealthy)+
  labs(
    title="Ratio of students enrolled in at least one AP course to the number of 11th and 12th graders.",
    subtitle="Philadelphia-Camden-Wilmington",
    x="% in AP courses",
    y="Frequency"
  )+
  theme_classic()+
  theme(panel.grid.major.x=element_line())
  
Philly_ap
```

From what we saw from analysis 4, we should not be too surprised that students from households who earn an income of at least 90K tend to be more ambitious in the classroom. This could be due to a number of factors. Students of lower incomes have a higher chance of having more responsibilities at home, such as taking care of younger siblings or working a part-time job. 



# Analysis 5: How each city compares in median income
data manip: 
```{r}
A <- data%>%
  drop_na(SE_MHE)

main_four <- A %>%
  filter (msaname15 %in% c("Los Angeles-Long Beach-Anaheim, CA Metro Area" ,"Philadelphia-Camden-Wilmington, PA-NJ-DE-MD Metro Area","Houston-The Woodlands-Sugar Land, TX Metro Area", "New York-Newark-Jersey City, NY-NJ-PA Metro Area" ))%>%
  group_by(msaname15)%>%
  summarise(mean_inc = mean(SE_MHE))%>%
  mutate(deviation= mean_inc - 71220.56)%>%
  mutate(region = sapply(strsplit(msaname15, "-"),"[", 1))%>%
  arrange(desc(mean_inc))

main_four%>%
   summarise(mean(mean_inc))
```

plot
```{r}
ggplot(main_four, aes(x = reorder(region, deviation), y = deviation,
           fill = deviation >0))+
  geom_bar(stat = "identity")+
  coord_flip()+
  scale_fill_discrete(name = "", labels = c("Below Average", "Above Average"))+
  labs(
    title="Household income",
    subtitle="We can see how each city is above or below the state average, $71,224",
    x = "City",
    y = "Amount Difference"
  )+
  theme_bw()

#How do I drop the "CA, Metro Area"
```

This graph is not interesting, just for reference. So all 4 cities make more than the national average, 60K. This is information tho that we can pull from the internet, it will be more accurate anyways. We can see if there exists a correlation between amount of income diff and %change in ap classes? 



