---
title: "Final Graph"
author: "Soyeon Park"
date: "12/3/2020"
output: github_document
---

# Necessary Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(gganimate)
library(maps)

library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE)
```

## Basic importating
```{r message=FALSE, warning=FALSE}
# Import data
df <- read_csv('data.csv')
```

# Data Cleaning

## Column names

According to the data dictionary, the unit of the 4 columns("ED_PRXECE", "ED_PRXHQECE", "HE_SUPRFND", "HE_RSEI") in the data is natural log. Natural log units change values less than e(2.71828) into negative numbers and this makes analysis difficult. Therefore, we are going to change the value of these 4 columns into original unit by using exp function.

```{r}
df <- df %>% 
  mutate(ED_PRXECE_NM = exp(ED_PRXECE)) %>%
  mutate(ED_PRXHQECE_NM = exp(ED_PRXHQECE)) %>%
  mutate(HE_SUPRFND_NM = exp(HE_SUPRFND)) %>%
  mutate(HE_RSEI_NM = exp(HE_RSEI)) %>%
  select(-ED_PRXECE, -ED_PRXHQECE, -HE_SUPRFND, -HE_RSEI)
```

Most of the columns are hard to understand. We tried to make it easier and obvious even without the data dictionary.

```{r}
df <- df %>% 
  rename(id = `_id`) %>% 
  rename(geo_id = geoid) %>% 
  rename(metro_areas = in100) %>%
  rename(area_code = msaid15) %>% 
  rename(county_code = countyfips) %>%
  rename(num_under_18 = pop) %>% 
  rename(ratio_students_AP_enrolled = ED_APENR) %>%
  rename(perc_over24_college_degree = ED_ATTAIN) %>%
  rename(perc_18to24_nearby_college_enrolled = ED_COLLEGE) %>%
  rename(perc_3to4_school_enrolled = ED_ECENROL) %>%
  rename(perc_high_grad = ED_HSGRAD) %>%
  rename(score_third_grade_math = ED_MATH) %>%
  rename(score_third_grade_read = ED_READING) %>%
  rename(perc_elementary_school_poverty = ED_SCHPOV) %>% 
  rename(perc_teacher_1and2_years=ED_TEACHXP) %>% 
  rename(num_ECE_nearby = ED_PRXECE_NM) %>% 
  rename(num_high_qual_ECE_nearby = ED_PRXHQECE_NM) %>% 
  rename(perc_supermarket_nearby = HE_FOOD) %>% 
  rename(perc_green_space_access = HE_GREEN) %>%
  rename(days_temp_above90 = HE_HEAT) %>%
  rename(perc_0to64_health_insurance = HE_HLTHINS) %>%
  rename(mean_ozone_amount = HE_OZONE) %>%
  rename(mean_microparticle = HE_PM25) %>%
  rename(perc_housing_vacancy = HE_VACANCY) %>%
  rename(index_walkability = HE_WALK) %>% 
  rename(num_waste_dump_sites = HE_SUPRFND_NM) %>%
  rename(index_air_pollutants = HE_RSEI_NM) %>%
  rename(perc_below100_poverty = SE_POVRATE) %>%
  rename(perc_household_public_assistance = SE_PUBLIC) %>%
  rename(perc_home_ownership = SE_HOME) %>%
  rename(perc_over15_high_skill = SE_OCC) %>%
  rename(median_income=SE_MHE) %>%
  rename(perc_adults_employed = SE_EMPRAT) %>%
  rename(perc_worker_commute_over1hour = SE_JOBPROX) %>%
  rename(perc_single_parent = SE_SINGLE)
```

```{r}
df_2015 <- df %>% filter(year == 2015)
df_2010 <- df %>% filter(year == 2010)
```

```{r}
df_2015_3rd <- df_2015 %>% mutate(score_third_grade = score_third_grade_math + score_third_grade_read) %>% 
  group_by(stateusps) %>% 
  summarise(third_grade = mean(score_third_grade, na.rm = TRUE))

df_2010_3rd <- df_2010 %>% mutate(score_third_grade = score_third_grade_math + score_third_grade_read) %>% 
  group_by(stateusps) %>% 
  summarise(third_grade = mean(score_third_grade, na.rm = TRUE))
```

Deleted unnecessary columns (They have the same values.)

1) Texas

```{r}
df_tx <- df %>% filter((county_code == "48201") | (county_code == "48157")| (county_code == "48339")|(county_code == "48291")|(county_code == "48473"))
```

```{r}
df_tx <- df_tx %>% mutate(county_name = case_when(county_code == "48201" ~ "Harris", county_code == "48157" ~ "Fort Bend", county_code == "48339" ~ "Montgomery", county_code == "48291" ~ "Liberty", county_code == "48473" ~ "Waller"))
```

```{r}
df_tx <- df_tx %>% mutate(group = ifelse(county_code == "48201", "city", "suburb")) %>%
  select(group, everything())
```

```{r}
df_tx$county_name = factor(df_tx$county_name, levels = c("Harris", "Fort Bend", "Montgomery", "Waller", "Liberty"))
```

2) Philadelphia
```{r}
df_pa <- df %>% filter((county_code == "42017")|(county_code == "42029")|(county_code == "42045")|(county_code == "42091")|(county_code == "42101"))
```

```{r}
df_pa <- df_pa %>% mutate(county_name = case_when(county_code == "42017" ~ "Bucks", county_code == "42029" ~ "Chester", county_code == "42045" ~ "Delaware", county_code == "42091" ~ "Montgomery", county_code == "42101" ~ "Philadelphia"))
```

```{r}
df_pa <- df_pa %>% mutate(group = ifelse(county_code == "42101", "city", "suburb")) %>%
  select(group, everything())
```

```{r}
df_pa$county_name = factor(df_pa$county_name, levels = c("Philadelphia", "Bucks", "Chester", "Delaware", "Montgomery"))
```

3) NY
```{r}
df_nyc <- df %>% filter((county_code == "36061") | (county_code == "36047")| (county_code == "36081")|(county_code == "36005")|(county_code == "36085"))
```

```{r}
df_nyc <- df_nyc %>% mutate(county_name = case_when(county_code == "36061" ~ "Manhattan", county_code == "36047" ~ "Brooklyn", county_code == "36081" ~ "Queens", county_code == "36005" ~ "Bronx", county_code == "36085" ~ "Staten Island"))
```

```{r}
df_nyc <- df_nyc %>% mutate(group = ifelse(county_code == "36061", "city", "suburb")) %>%
  select(group, everything())
```

```{r}
df_nyc$county_name = factor(df_nyc$county_name, levels = c("Manhattan", "Bronx", "Brooklyn", "Queens", "Staten Island"))
```

4) CA
```{r}
df_ca <- df %>% filter((county_code == "06111") | (county_code == "06059")| (county_code == "06029")|(county_code == "06071")|(county_code == "06037"))
```

```{r}
df_ca <- df_ca %>% 
  mutate(county_name = case_when(county_code == "06111" ~ "Ventura", 
                                county_code == "06059" ~ "Orange", 
                                county_code == "06029" ~ "Kern", 
                                county_code == "06071" ~ "San Bernardino", 
                                county_code == "06037" ~ "Los Angeles"))
```

```{r}
df_ca <- df_ca %>% mutate(group = ifelse(county_code == "06037", "city", "suburb")) %>%
  select(group, everything())
```

```{r}
df_ca$county_name = factor(df_ca$county_name, levels = c("Los Angeles", "Ventura", "Orange", "Kern",  "San Bernardino"))
```


```{r}
df_tx_2010 <- df_tx %>% filter(year == "2010")
df_tx_2015 <- df_tx %>% filter(year == "2015")
```

```{r}
df_pa_2010 <- df_pa %>% filter(year == "2010")
df_pa_2015 <- df_pa %>% filter(year == "2015")
```

```{r}
df_nyc_2010 <- df_nyc %>% filter(year == "2010")
df_nyc_2015 <- df_nyc %>% filter(year == "2015")
```

```{r}
df_ca_2010 <- df_ca %>% filter(year == "2010")
df_ca_2015 <- df_ca %>% filter(year == "2015")
```

```{r}
df_tx_2015 %>% 
  mutate(Area = ifelse(county_name == "Harris", "City", "Suburb")) %>% 
  ggplot(aes(x = county_name, y = score_third_grade_math + score_third_grade_read, fill = Area, alpha = Area)) +
  geom_boxplot() +
  geom_hline(data = df_2015, aes(yintercept = mean(score_third_grade_read, na.rm = TRUE) + mean(score_third_grade_math, na.rm = TRUE)), col = "red",linetype='dotted') + 
  scale_fill_manual(values = c("#69b3a2", "grey")) +
  scale_alpha_manual(values=c(1,0.1)) +
  theme_classic() +
  labs(title = "Houston 3rd Grades Academic Scores by County", subtitle = "The red lines show U.S nation's average", x = "County", y = "3rd Grades Academic score")
```
```{r}
df_pa_2015 %>% 
  mutate(Area = ifelse(county_name == "Philadelphia", "City", "Suburb")) %>% 
  ggplot(aes(x = county_name, y = score_third_grade_math + score_third_grade_read, fill = Area, alpha = Area)) +
  geom_boxplot(outlier.shape = NA) +
  geom_hline(data = df_2015, aes(yintercept = mean(score_third_grade_read, na.rm = TRUE) + mean(score_third_grade_math, na.rm = TRUE)), col = "red",linetype='dotted') + 
  scale_fill_manual(values = c("#69b3a2", "grey")) +
  scale_alpha_manual(values=c(1,0.1)) +
  theme_classic() +
  labs(title = "Philadelphia 3rd Grades Academic Scores by County", subtitle = "The red lines show U.S nation's average", x = "County", y = "3rd Grades Academic score")
```

```{r}
df_ca_2015 %>% 
  mutate(Area = ifelse(county_name == "Los Angeles", "City", "Suburb")) %>% 
  ggplot(aes(x = county_name, y = score_third_grade_math + score_third_grade_read, fill = Area, alpha = Area)) +
  geom_boxplot(outlier.shape = NA) +
  geom_hline(data = df_2015, aes(yintercept = mean(score_third_grade_read, na.rm = TRUE) + mean(score_third_grade_math, na.rm = TRUE)), col = "red",linetype='dotted') + 
  scale_fill_manual(values = c("#69b3a2", "grey")) +
  scale_alpha_manual(values=c(1,0.1)) +
  theme_classic() +
  labs(title = "Los Angeles 3rd Grades Academic Scores by County", subtitle = "The red lines show U.S nation's average", x = "County", y = "3rd Grades Academic score")
```
```{r}
df_nyc_2015 %>% 
  mutate(Area = ifelse(county_name == "Manhattan", "City", "Suburb")) %>% 
  ggplot(aes(x = county_name, y = score_third_grade_math + score_third_grade_read, fill = Area, alpha = Area)) +
  geom_boxplot(outlier.shape = NA) +
  geom_hline(data = df_2015, aes(yintercept = mean(score_third_grade_read, na.rm = TRUE) + mean(score_third_grade_math, na.rm = TRUE)), col = "red",linetype='dotted') + 
  scale_fill_manual(values = c("#69b3a2", "grey")) +
  scale_alpha_manual(values=c(1,0.1)) +
  theme_classic() +
  labs(title = "New York City 3rd Grades Academic Scores by County", subtitle = "The red lines show U.S nation's average", x = "County", y = "3rd Grades Academic score")
```

```{r}
df_tx_2015_resi <- df_tx_2015 %>% filter(!is.na(perc_supermarket_nearby)) %>%
  filter(!is.na(perc_green_space_access)) %>%
  filter(!is.na(index_walkability)) %>%
  mutate(residential_environment = scale(perc_supermarket_nearby) + scale(perc_green_space_access) + scale(index_walkability)) %>%
  group_by(county_name) %>%
  summarise(resi_mean = mean(residential_environment))

ggplot(df_tx_2015_resi, aes(x = reorder(county_name, resi_mean, desc), y = resi_mean, fill=factor(ifelse(county_name=="Harris","City","Suburb")))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(name = "Area", values=c("red","grey50")) +
  theme_bw() + 
  labs(title = "Houston Index of good residential environment by conuty", subtitle = "Residential environment = Near Supermarket + Green Space Access + Walkability", y = "Index of good residential environment", x = "County" ) 
```

```{r}
df_pa_2015_resi <- df_pa_2015 %>% filter(!is.na(perc_supermarket_nearby)) %>%
  filter(!is.na(perc_green_space_access)) %>%
  filter(!is.na(index_walkability)) %>%
  mutate(residential_environment = scale(perc_supermarket_nearby) + scale(perc_green_space_access) + scale(index_walkability)) %>%
  group_by(county_name) %>%
  summarise(resi_mean = mean(residential_environment))

ggplot(df_pa_2015_resi, aes(x = reorder(county_name, resi_mean, desc), y = resi_mean, fill=factor(ifelse(county_name=="Philadelphia","City","Suburb")))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(name = "Area", values=c("red","grey50")) +
  theme_bw() + 
  labs(title = "Philadelphia Index of good residential environment by conuty", subtitle = "Residential environment = Near Supermarket + Green Space Access + Walkability", y = "Index of good residential environment", x = "County" )
```

```{r}
df_nyc_2015_resi <- df_nyc_2015 %>% filter(!is.na(perc_supermarket_nearby)) %>%
  filter(!is.na(perc_green_space_access)) %>%
  filter(!is.na(index_walkability)) %>%
  mutate(residential_environment = scale(perc_supermarket_nearby) + scale(perc_green_space_access) + scale(index_walkability)) %>%
  group_by(county_name) %>%
  summarise(resi_mean = mean(residential_environment))

ggplot(df_nyc_2015_resi, aes(x = reorder(county_name, resi_mean, desc), y = resi_mean, fill=factor(ifelse(county_name=="Manhattan","City","Suburb")))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(name = "Area", values=c("red","grey50")) +
  theme_bw() + 
  labs(title = "NYC Index of good residential environment by conuty", subtitle = "Residential environment = Near Supermarket + Green Space Access + Walkability", y = "Index of good residential environment", x = "County" )
```

```{r}
df_ca_2015_resi <- df_ca_2015 %>% filter(!is.na(perc_supermarket_nearby)) %>%
  filter(!is.na(perc_green_space_access)) %>%
  filter(!is.na(index_walkability)) %>%
  mutate(residential_environment = scale(perc_supermarket_nearby) + scale(perc_green_space_access) + scale(index_walkability)) %>%
  group_by(county_name) %>%
  summarise(resi_mean = mean(residential_environment))

ggplot(df_ca_2015_resi, aes(x = reorder(county_name, resi_mean, desc), y = resi_mean, fill=factor(ifelse(county_name=="Los Angeles","City","Suburb")))) +
  geom_bar(stat = "identity") +
  theme_bw() +
  scale_fill_manual(name = "Area", values=c("red","grey50")) +
  labs(title = "LA Index of good residential environment by conuty", subtitle = "Residential environment = Near Supermarket + Green Space Access + Walkability", y = "Index of good residential environment", x = "County" )
```
