---
title: "NYC Economics"
---

## Economic Variablity in the City
```{r}
df_nyc <- df %>% filter(
    county_code == '36061' |
    county_code ==  '36047' |
    county_code ==  '36081' |
    county_code ==  '36005' |
    county_code ==  '36085' 
)

df_nyc <- df_nyc %>%
  mutate(county_code = replace(county_code, county_code == '36061', 'Manhattan')) %>% 
  mutate(county_code = replace(county_code, county_code == '36047', 'Brooklyn')) %>%
  mutate(county_code =  replace(county_code, county_code == '36081', 'Queens')) %>%
  mutate(county_code =  replace(county_code, county_code == '36005', 'Bronx')) %>%
  mutate(county_code =  replace(county_code, county_code == '36085', 'Staten Island'))%>%
  mutate_all(~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))
```

```{r}
df_adult_success <- df_nyc %>% select(county_code,college_deg,median_income,phealth_insurance,perc_over15_high_skill,home_ownership,employment_rate)
```

### Median Income
```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = county_code, y = median_income))+
  geom_violin() + 
  geom_jitter(aes(y = median_income), alpha = .2)
```
The first economic indicator of economic success seems pretty obvious - the amount of money made in terms of the median. The mean is avoided here to prevent extreme cases from driving up or down the amount. From the graph the first two that stick out are Manhattan and the Bronx for the opposite reasons. Manhattan seems to have more economic success and the Bronx seems to be struggling. As the financial center of the world it does not come as a surprise to Manhattans success. However, given the proximity of the Bronx to Manhattan the reason for the Bronx's struggle is not as clear.

```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = college_deg, y = median_income))+
  geom_point(aes(color = county_code), alpha = .5)+
  scale_y_continuous(limits = c(0,150000))
```

The first factor that one could guess that would potentially drive economic success is college degrees(add citation). As such, it does not come as a surprise that there appears to be a nearly one to one relationship between the two. However, what is interesting is that on the first graph Staten Island seemed to come in second in terms of median income but appears to have less college degrees as a percentage than Brooklyn(look more into this). Unsurprisingly Manhattan seems to have the most college degrees and highest financial success.

```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = county_code, y = employment_rate))+
  geom_violin() + 
  geom_jitter(aes(y = employment_rate), alpha = .2)
```
What is interesting to note here is that there does not seem to be much in the way of difference in terms of the employment rate. However, this graph can be a bit misleading as everything seems to be high. 75% unemployment is a very large portion of the population. To put this into perspective, unemployment in the United States is generally around 90%. Staten Island seems like the outliers here exhibiting a high income  but a lower rate of college degrees and employment (examine more).



```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = college_deg, y = employment_rate))+
  geom_point(aes(color = county_code), alpha = .5)
```
This is one of the most interesting graphs. It appears from this graph that the employment rate is nearly devoid of a relationship with a college degree expect in the extreme cases of a lack of degree. However, it is very possible that relationship is actually inverted with the employment generating whether someone has the opportunity to go to college instead of the other way around.  

 

```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = perc_over15_high_skill, y = employment_rate))+
  geom_point(aes(color = county_code), alpha = .5)
```
Given the relationship above it seems to make sense that perhaps is being a high skilled worked that drives employment rate. However, that also does not seem to be true expect again in the extreme case. 
```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = perc_over15_high_skill, y = median_income))+
  geom_point(aes(color = county_code), alpha = .5)
```
However, college degrees are not the sole measure for financial stability. There are numerous other factors like being high skilled worker which would include things like trade school. Here you do see a bit more of a relationship between high skill and median income though not one as stark as having a college degree. 

```{r}
df_adult_success %>% group_by(county_code) %>% ggplot(aes(x = county_code, y = home_ownership))+
  geom_violin() + 
  geom_jitter(aes(y = home_ownership, color = median_income), alpha = .5)+
  scale_color_viridis()
```
The last graph might seem like it belongs in livability which is an argument is  fairly solid. However, given that often times a house seems to be the largest asset housing rates are also right at home in financial success. While one might expect Manhattan to have the largest amount of homes it is apparent this is not the case until you progress into the super wealthy. Given the price of real estate in the city this makes sense. The Bronx also matchs what you would expect as does Staten Island as real estate prices drop. 